### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@cae9ebd294a5cdb86e6726c5c2b15722c29476a8 🚀

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence. In a previous
work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
of model performance than the True Skill Statistic (TSS), which we will
refer to as Youden's informedness (or \(J\)); these conclusions were
later criticized by Somodi et al. (2017), who emphasized that
informedness' relationship to prevalence depends on assumptions about
bias in the model, and therefore recommend the use of \(\kappa\) as a
validation of classification performance. Although this work offers
recommendations about the comparison of models, it doesn't establishes
baselines or good practices for training on imbalanced ecological data.
Within the context of networks, there are three specific issues that
need to be adressed. First, what values of performance measures are we
expecting for a classifier that has poor performance? This is
particularly important as it can evaluate whether low prevalence can
lull us into a false sense of predictive accuracy. Second, independently
of the question of model evaluation, is low prevalence an issue for
\emph{training}, and can we remedy it? Finally, because the low amount
of data on interaction makes a lot of imbalance correction methods (see
\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
optimized with the least amount of positive interaction data?

@@

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, amd can therefore return
somethins considered to be a pseudo-probability); in this case, the
optimal value below which predicitions are assumed to be negative
(\emph{i.e.} the interaction does not exist) can be determined by
picking a threshold maximizing some value on the ROC curve or the PR
curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
(2015) established that the ROC-AUC is biased towards over-estimating
performance for imbalanced data; on the contrary, the PR-AUC is able to
identify classifiers that are less able to detect positive interactions
correctly, with the additional advantage of having a baseline value
equal to prevalence. Therefore, it is important to assess whether these
two measures return different results when applied to ecological network
prediction.

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidental correct predictions
of negatives. The \(F_1\) score and positive predictive values are less
sensitive to bias, but \textbf{TODO}

@@

\hypertarget{definition-of-the-performance-measures}{%
\subsection{Definition of the performance
measures}\label{definition-of-the-performance-measures}}

\(\kappa\)

informedness

MCC

ROC-AUC

\(F_{\beta}\)

PR-AUC - baseline is prevalence, wtf are recall and precision, you may
ask?

@@

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}rccccccc@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
three network structure metrics, for 250 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endfirsthead
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endhead
Decision tree & 0.85 & 0.92 & 0.97 & 0.12 & 0.21 & 0.76 &
0.31\tabularnewline
BRT & 0.90 & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 & 0.27\tabularnewline
Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
\textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
Ridge Regression & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 &
0.18\tabularnewline
Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
0.75 & 0.31\tabularnewline
Data & & & & & 0.18 & 0.66 & 0.34\tabularnewline
\bottomrule
\end{longtable}

@@

The results presented here highlight an interesting paradox: larger
networks (with lower connectance) require more training set bias in
order to maximize model performance fig.~\ref{fig:optimbias}, but are
also more difficult to predict according to MCC and PR-AUC
fig.~\ref{fig:optimperf}. This suggests that the task of network
prediction will be difficult regardless of network size: by being
limited by the \emph{frequency} of interactions when the network is
large, and by being limited by the \emph{number} of interactions when
the network is small. Nevertheless, based on the simulations and
numerical experiments, it is possible to formulate a series of
recommendations for the evaluation of network prediction models.

@@

Second, because the PR-AUC responds more to network connectance
fig.~\ref{fig:optimperf} and training set imbalance
fig.~\ref{fig:biasrocpr}, it should be used as a measure of model
performance over the ROC-AUC. This is not to say that ROC-AUC should be
discarded (in fact, a low ROC-AUC is a sign of an issue with the model),
but that its interpretation should be guided by the PR-AUC value. This
again echoes recommendations from other fields (Saito \& Rehmsmeier,
2015).

Thirdly, regardless of network connectance \emph{or} measure to evaluate
the model performance, as long as the network connectance is larger than
\(\approx 0.1\), artificially balancing the training set to have
equiprevalence will give the best possible results. This was true for
all models.

Finally, it is noteworthy that the ensemble model was systematically
better than the component models; even when poor models were included
(Random Forest and Decision Tree), the ensemble was able to leverage the
different biases expressed by the models to make an overall more
accurate prediction. We do not expect that ensembles will \emph{always}
be better than single models. In a recent multi-model comparison, Becker
et al. (2021) found that the ensemble was \emph{not} the best model.
There is no general conclusion to draw from this besides reinforcing the
need to be pragmatic about which models should be included in the
ensemble, or whether to use an ensemble at all. In a sense, the
surprising peformance of the ensemble model should form the basis of the
last recommendation: optimal training set bias and its interaction with
connectance and binary classifier is, in a sense, an hyperparameter that
should be assessed. The distribution of results in
fig.~\ref{fig:optimbias} and fig.~\ref{fig:optimperf} show that there
are variations around the trend; furthermore, networks with different
structures than the one we simulated here may respond in different ways.

@@

@@

@@

>>>>

There is an immense amount of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013); three
threshold metrics (\(\kappa\), informedness, and the Matthews
Correlation Coefficient), and two ranking metrics (the ROC-AUC and
PR-AUC). The \(\kappa\) measure of agreement (Landis \& Koch, 1977)
establishes the extent to which two observers (here the data and the
prediction) agree, and is measured as

\[
2\frac{tp\times tn - fn\times fp}{(tp+fp)\times (fp+tn)+(tn+fp)\times (tn+fn)} \,.
\]

Informedness (\textbf{Youden1950IndRat?}) (also known as bookmaker
informedness or the True Skill Statistic) is
\(\text{TPR}+\text{TNR}-1\), where \(\text{TPR}= tp/(tp+fn)\) and
\(\text{TNR} = tn/(tn+fp)\); informedness can be used to find the
optimal cutpoint in thresholding analyses (Schisterman et al., 2005).
The MCC is defined as

\[
\frac{tp\times tn - fn\times fp}{\srt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
\]

Finally, \(F_1\) is the harmonic mean of precision (the chance that a
positive even was correctly classified) and sensitivity (the ability to
correctly classify positive events), and is defined as

\[
2\frac{tp}{2\times tp + fp + fn}\,.
\]

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, amd can therefore return
somethins considered to be a pseudo-probability); in this case, the
optimal value below which predicitions are assumed to be negative
(\emph{i.e.} the interaction does not exist) can be determined by
picking a threshold maximizing some value on the ROC curve or the PR
curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
(2015) established that the ROC-AUC is biased towards over-estimating
performance for imbalanced data; on the contrary, the PR-AUC is able to
identify classifiers that are less able to detect positive interactions
correctly, with the additional advantage of having a baseline value
equal to prevalence. Therefore, it is important to assess whether these
two measures return different results when applied to ecological network
prediction. The ROC curve is defined by the false positive rate on the
\(x\) axis, and the true positive rate on the \(y\) axis, and the PR
curve is defined by the true positive rate on the \(x\) axis, and the
positive predictive value on the \(y\) axis. By comparison with the
previous paragraph, it is obvious that \(F_1\) has ties to the PR curve
(being close to the expected PR-AUC), and that informedness has ties to
the ROC curve (whereby the threshold maximizing informedness is also the
point of maximal inflection on the ROC curve). One important difference
between ROC and PR is that the later does not prominently account for
the size of the true negative compartments: in short, it is more
sensitive to the correct positive predictions. In a context of strong
imbalance, PR-AUC is therefore a more stringent test of model
performance.

The same approach is used to evaluate \emph{e.g.} species distribution
models (SDMs). Indeed, the training and evaluation of SDMs as binary
classifiers suffers from the same issue of low prevalence. In previous
work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
of model performance than the True Skill Statistic (TSS; which we refer
to as Youden's informedness); these conclusions were later criticized by
Somodi et al. (2017), who emphasized that informedness' relationship to
prevalence depends on assumptions about bias in the model, and therefore
recommend the use of \(\kappa\) as a validation of classification
performance. Although this work offers recommendations about the
comparison of models, it doesn't establishes baselines or good practices
for training on imbalanced ecological data. Within the context of
networks, there are three specific issues that need to be adressed.
First, what values of performance measures are we expecting for a
classifier that has poor performance? This is particularly important as
it can evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training}, and can we
remedy it? Finally, because the low amount of data on interaction makes
a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
2015) hard to apply, which indicators can be optimized with the least
amount of positive interaction data?

@@

We establish that due to the low prevalence of interactions, even poor
classifiers applied to food web data will reach a high accuracy; this is
because the measure is dominated by the accidental correct predictions
of negatives. On simulated confusion matrices with ranges of imbalance
that are credible for ecological networks, MCC had the most desirable
behavior, and informedness is a linear measure of classifier skill. By
performing simulations with four models and an ensemble,we show that
informedness and ROC-AUC are consistently high on network data, and that
MCC and PR-AUC are more accurate measures of the effective performance
of the classifier. Finally, by measuring the structure of predicted
networks, we highlight an interesting paradox: the models with the best
performance measures are not the models with the closest reconstructed
network structure. We discuss these results in the context of
establishing guidelines for the prediction of ecological interactions.

@@

@@

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}rccccccc@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
three network structure metrics, for 250 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endfirsthead
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
\(Q\)\tabularnewline
\midrule
\endhead
Decision tree & 0.85 & 0.92 & 0.97 & 0.12 & 0.21 & 0.76 &
0.31\tabularnewline
BRT & \textbf{0.90} & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 &
0.27\tabularnewline
Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
\textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
Ridge Regression & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 &
0.18\tabularnewline
Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
0.75 & 0.31\tabularnewline
Data & & & & & 0.18 & 0.66 & 0.34\tabularnewline
\bottomrule
\end{longtable}

@@

The results presented here highlight an interesting paradox: although
the Random Forest was ultimately able to get a correct estimate of
network structure tbl.~\ref{tbl:comaprison}, it ultimately remains a
poor classifier, as evidenced by its low PR-AUC. This suggests that the
goal of predicting \emph{interactions} and predicting \emph{networks}
may not be solvable in the same way -- of course a perfect classifier of
interactions would make a perfect network prediction; but even the best
scoring predictor of interactions (the ensemble model) had not
necessarilly the best prediction of network structure. The tasks of
predicting networks structure and of predicting interactions within
networks are essentially two different ones. For some applications
(\emph{.e.g.} comparison of network structure across gradients), one may
care more about a robust estimate of the structure, at the cost at
putting some interactions at the wrong place. For other applications
(\emph{e.g.} identifying pairs of interacting species), one may
conversely care more about getting as many pairs right, even though the
mistakes accumulate in the form of a slightly worse estimate of network
structure. How these two approaches can be reconciled is undoubtedly a
task for further research. Despite this apprent tension at the heart of
the predictive exercise, we can use the results presented here to
suggest a number of guidelines.

@@

Second, because the PR-AUC responds more to network connectance
fig.~\ref{fig:optimperf} and training set imbalance
fig.~\ref{fig:biasrocpr}, it should be used as a measure of model
performance over the ROC-AUC. This is not to say that ROC-AUC should be
discarded (in fact, a low ROC-AUC is a sign of an issue with the model),
but that its interpretation should be guided by the PR-AUC value.
Specifically, a high ROC-AUC is not informative, as it can be associated
to a low PR-AUC (see \emph{e.g.} Random Forest in
tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015).

Thirdly, regardless of network connectance, maximizing informedness
required a training set bias of about 0.5, and maximizing the MCC
required a training set bias of 0.7 and more. This has an important
consequence in ecological networks, for which the pool of positive cases
(interactions) to draw from is typically small: the most parsimonious
measure (\emph{i.e.} the one requiring to discard the least amount of
information to train the model) will give the best validation potential,
and is probably the informedness (maximizing informedness is the
generally accepted default for imbalanced classification; Schisterman et
al., 2005).

Finally, it is noteworthy that the ensemble model was systematically
better than the component models; even when the models were individually
far form perfect, the ensemble was able to leverage the different biases
expressed by the models to make an overall more accurate prediction. We
do not expect that ensembles will \emph{always} be better than single
models. In a recent multi-model comparison, Becker et al. (2021) found
that the ensemble was \emph{not} the best model. There is no general
conclusion to draw from this besides reinforcing the need to be
pragmatic about which models should be included in the ensemble, or
whether to use an ensemble at all. In a sense, the surprising peformance
of the ensemble model should form the basis of the last recommendation:
optimal training set bias and its interaction with connectance and
binary classifier is, in a sense, an hyperparameter that should be
assessed. The distribution of results in fig.~\ref{fig:optimbias} and
fig.~\ref{fig:optimperf} show that there are variations around the
trend; furthermore, networks with different structures than the one we
simulated here may respond in different ways.

@@

\leavevmode\hypertarget{ref-Ferri2009ExpCom}{}%
Ferri, C., Hernández-Orallo, J., \& Modroiu, R. (2009). An experimental
comparison of performance measures for classification. \emph{Pattern
Recognition Letters}, \emph{30}(1), 27--38.
\url{https://doi.org/10.1016/j.patrec.2008.08.010}

\leavevmode\hypertarget{ref-He2013ImbLea}{}%
He, H., \& Ma, Y. (Eds.). (2013). \emph{Imbalanced Learning:
Foundations, Algorithms, and Applications} (1st edition). Wiley-IEEE
Press.

@@

\leavevmode\hypertarget{ref-Landis1977MeaObs}{}%
Landis, J. R., \& Koch, G. G. (1977). The Measurement of Observer
Agreement for Categorical Data. \emph{Biometrics}, \emph{33}(1),
159--174. \url{https://doi.org/10.2307/2529310}

@@

\leavevmode\hypertarget{ref-Schisterman2005OptCut}{}%
Schisterman, E. F., Perkins, N. J., Liu, A., \& Bondell, H. (2005).
Optimal Cut-point and Its Corresponding Youden Index to Discriminate
Individuals Using Pooled Blood Samples. \emph{Epidemiology},
\emph{16}(1), 73--81.
\url{https://doi.org/10.1097/01.ede.0000147512.81966.ba}
